---
output: html_document
---

Reproducible Research Course Project 1
======================================

In this assignment, data from a personal activity monitoring device will be analysed to answer the given questions on the number of steps per day and the average daily activity pattern.


### Loading the data
```{r load}
activity<-read.csv("activity.csv")
```

### Steps taken per day
```{r stepsperday}
completeactivity<-activity[complete.cases(activity),]
stepsperday<-tapply(completeactivity$steps, completeactivity$date, sum)

hist(stepsperday, main="Total number of steps per day", xlab="Total number of steps per day", col="grey")

mu<-mean(stepsperday,na.rm=T)
me<-median(stepsperday,na.rm=T)

options(scipen=999)
```

The mean number of total steps taken per day is `r mu`, the median number of total steps taken per day is `r me`.

### Average daily activity pattern
```{r dailyactivity}
stepsperinterval<-tapply(completeactivity$steps, 
      completeactivity$interval, mean)
plot(rownames(stepsperinterval), stepsperinterval, type="l",                   
      main="Average daily activity", 
      xlab="Interval", ylab="Average number of steps")

whichmax<-names(which.max(stepsperinterval))
max<-stepsperinterval[rownames(stepsperinterval)==whichmax]

abline(v=whichmax, col="red")
```

Interval `r whichmax` contains on average the maximum number of steps (`r round(max,2)` steps).

### Imputing missing values
```{r missingimputation}
nmiss<-sum(sapply(activity$steps, FUN=is.na))

imputedactivity<-activity

intervals<-unique(activity$interval)

for (i in 1:length(stepsperinterval)){
      imputedactivity$steps[imputedactivity$interval==intervals[i] & is.na(imputedactivity$steps)]<-stepsperinterval[rownames(stepsperinterval)==intervals[i]]
}

nmiss2<-sum(sapply(imputedactivity$steps, FUN=is.na))
```

The total number of rows with NA's is `r nmiss`. The missing numbers of steps are imputed by the mean number of steps per interval. Thus, in the new dataset `imputedactivity`, there are no missing numbers of steps left (`nmiss2`=`r nmiss2`).

### Steps taken per day (2)
```{r stepsperday2}
stepsperday2<-tapply(imputedactivity$steps, imputedactivity$date, sum)

hist(stepsperday2, main="Total number of steps per day (based on imputed data)", xlab="Total number of steps per day", col="grey")

mu2<-mean(stepsperday2,na.rm=T)
me2<-median(stepsperday,na.rm=T)
```

The mean number of total steps taken per day, based on the imputed data, is `r mu2`, the median number of total steps taken per day is `r me2`. The difference based on the original data versus the imputed data is `r mu-mu2` and `r me-me2`, respectively.

### Daily activity pattern on weekdays vs weekenddays
```{r weekvsweekend}
imputedactivity$weekday<-weekdays(as.Date(imputedactivity$date,
                                          format="%Y-%m-%d"))
imputedactivity$weekend<-as.factor(ifelse(imputedactivity$weekday=="samedi"|
                                      imputedactivity$weekday=="dimanche",
                                "weekend","weekday"))
imputedactivityWE<-imputedactivity[imputedactivity$weekend=="weekend",]
imputedactivityWD<-imputedactivity[imputedactivity$weekend=="weekday",]

stepsperintervalWE<-tapply(imputedactivityWE$steps, 
      imputedactivityWE$interval, mean)
stepsperintervalWD<-tapply(imputedactivityWD$steps, 
      imputedactivityWD$interval, mean)

par(mfrow=c(1,2))
plot(rownames(stepsperintervalWE), stepsperintervalWE,
     type="l",                   
     main="Average daily activity (weekends)", 
     xlab="Interval", ylab="Average number of steps",
     ylim=c(0,250),
     col="blue")
plot(rownames(stepsperintervalWD), stepsperintervalWD,
     type="l",                   
     main="Average daily activity (weekdays)", 
     xlab="Interval", ylab="Average number of steps",
     ylim=c(0,250),
     col="red")
```

Based on the daily activity plots for weekenddays (left) and weekdays (right), it is clear that the activity patterns are quite different. On weekdays, there is much activity is a sudden increase in activity in the morning, and not much later a peak in activity, which might result from rushing to work; then all day is low in activity, with a slight increase again at the end of the workday. On weekenddays, the average activity is less extreme.